---
title: <center><strong>Data Prep Template</strong></center>
date: <center>23/11/2019</center>
output:
  html_document: 
    toc: yes
    toc_depth: 6
---

These document contains functions that I use for data prep at work.
<br>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```


## 0. Specify all the functions to be used in this analysis
<br>
```{r, echo =FALSE, warning=FALSE, message=FALSE}

### 1. package_func: A function for installing packages

package_func <- function(pkgs){

miss_pkgs <- pkgs[!pkgs %in% installed.packages()[,1]]

## Install the missing packages

if(length(miss_pkgs)>0){
  install.packages(miss_pkgs)
}

## Loading all the packages

invisible(lapply(pkgs,library,character.only=TRUE))

## Remove the objects that are no longer required

rm(miss_pkgs)
rm(pkgs)

}
# e.g.
## pkgs <- c("dplyr","lubridate","tidyr","ggplot2","kableExtra","openxlsx","readxl","janitor","data.table","haven","stringr","DT")
## package_func(pkgs)


### 2. factor_func: A function for factoring variables
factor_func <- function(data, var, levs) {
  
data[,var] <- factor(data[,var],
                    levels = levs,
                    labels = levs) 
return(data[,var])
}

#e.g.
## df[,"Gender"]<- factor_func(df, "Gender", c("Male","Female"))
## df[,"Role"]<- factor_func(df, "Role", c("Junior","Mid", "Senior"))

### 3. sata_factor_func: A function for factoring select all variables

pattern <- "specify pattern here"
sata_vars <- grep(pattern, df, ignore.case = T, value = T)
for(i in 1: length(sata_vars)){
df[,sata_vars[i]] <- factor_func(df, sata_vars[i], levs)
}

```

## 0. Install and load all the packages
<br>
```{r, echo =FALSE, warning=FALSE, message=FALSE}

pkgs <- c("dplyr","lubridate","tidyr","ggplot2","kableExtra","openxlsx","readxl","janitor","data.table","haven","stringr","DT","sjlabelled")

package_func(pkgs)
```




## 1. Read in the dataset
<br>

```{r, echo =FALSE, warning=FALSE, message=FALSE}

df <- read_xlsx("")

```

## 2. Covert "N/A"s to system N/As
<br>
```{r, echo =FALSE, warning=FALSE, message=FALSE}


```

## 3. Remove system variables
<br>
```{r, echo =FALSE, warning=FALSE, message=FALSE}


```


## 4. Remove duplicates
<br>
```{r, echo =FALSE, warning=FALSE, message=FALSE}


```


## 5. Drop test cases
<br>
```{r, echo =FALSE, warning=FALSE, message=FALSE}


```

## 6. Check that screeners worked
<br>
```{r, echo =FALSE, warning=FALSE, message=FALSE}


```

## 7. Check that logics worked
<br>
```{r, echo =FALSE, warning=FALSE, message=FALSE}


```

## 8. Factorise the single choice questions
<br>
```{r, echo =FALSE, warning=FALSE, message=FALSE}
## df[,"Gender"]<- factor_func(df, "Gender", c("Male","Female"))

```

## 9. Factorise the "select all that apply" questions
<br>
```{r, echo =FALSE, warning=FALSE, message=FALSE}
## df[,"Gender"]<- factor_func(df, "Gender", c("Male","Female"))

```

## 10. Cut back data

<br>
```{r, echo =FALSE, warning=FALSE, message=FALSE}

```

## 11. Code open ended questions

<br>
```{r, echo =FALSE, warning=FALSE, message=FALSE}

```

## 12. Organise the variables in the way they should appear in the data 
<br>
```{r, echo =FALSE, warning=FALSE, message=FALSE}

```

## 13. Attach variable labels, this will make work easier and faster in SPSS
<br>
```{r, echo =FALSE, warning=FALSE, message=FALSE}
vars <- colnames(df)

labs <- c("insert the variable labels here, as they appear in the subform")

varlabel_df <- data.frame(vars, labs) ##This is used to just check that the variables and variable labels match

df <- set_label(df, label = labs)

```


## 14. Save the data in the internal folder, as .xlsx and .sav
<br>
```{r, echo =FALSE, warning=FALSE, message=FALSE}

```
