---
title: <center><strong>Trivago Case Study</strong></center>
subtitle: <center><strong><h2>Challenge 428 - Building a predictive 'hits' model</h2></strong></center>
author: <center><strong><h1>[Shelmith Nyagathiri Kariuki](https://github.com/Shelmith-Kariuki)</h1></strong></center>
date: <center><strong><h2>January 20, 2020</h2></strong></center>
output:
  html_document:
    toc: yes
    toc_depth: '6'
    code_folding: hide
---
Trivago is an online company that meets the growing demand for independent travel information. It offers an extensive hotel meta search to travellers. 

Trivago hosts different case studies on their website, that job applicants can work on.

Details about this particular can be found can be found [here](https://jobs.zealpath.com/m/case/detail/428). 

The variables of the dataset are defined below:

+ row_num: a number uniquely identifying each row.

+ locale: the platform of the session.

+ day_of_week: Mon-Fri, the day of the week of the session.

+ hour_of_day: 00-23, the hour of the day of the session.

+ agent_id: the device used for the session.

+ entry_page: describes the landing page of the session.

+ path_id_set: shows all the locations that were visited during the session.

+ traffic_type: indicates the channel the user cane through eg. search engine, email, ...

+ session_duration: the duration in seconds of the session.

+ hits: the number of interactions with the trivago page during the session.


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE,fig.width = 12)

```

### 0. Install the libraries required
<br>
```{r,warning=FALSE, message=FALSE}

## create a vector of packages to be installed
pkgs <- c("tidyverse","data.table","DT","lubridate","ggthemes","randomForest")

## Check if there are packages you want to load, that are not already installed. 
miss_pkgs <- pkgs[!pkgs %in% installed.packages()[,1]] 

## Installing the missing packages
if(length(miss_pkgs)>0){
  install.packages(miss_pkgs)
}
## Loading all the packages
invisible(lapply(pkgs,library,character.only=TRUE))

## Remove the objects that are no longer required
rm(miss_pkgs)
rm(pkgs)

```


<br>
```{r,warning=FALSE, message=FALSE}
### Setting the plot theme
trivago_theme<- theme_hc()+ theme(legend.position = "right", 
                 legend.direction = "vertical",
                 legend.title = element_blank(),
                 plot.title = element_text( size = rel(1.6), hjust = 0.5),
                plot.subtitle = element_text(size = rel(1.5), hjust = 0.5),
                 #axis.text = element_text( size = rel(1.5)),
                 axis.text.x = element_text(size =rel(1.5),angle = 0),
                 axis.text.y = element_text(size =rel(1.5),angle = 0),
                 axis.title = element_text( size = rel(1.55)),
                axis.line.x = element_line(size = 1.5, colour = "#c94a38"),
                 panel.background = element_rect(fill = NA))

### Colours that will be used for the plots 
trivago_blue = "#007faf"
trivago_orange = "#c94a38"
trivago_yellow = "#f48f00"

## Avoidance of scientific numbers
options(scipen = 999)

## Printing function
pr_func<-function(data,cnames){
datatable(data,colnames = cnames,
  extensions = 'Buttons', options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'print')
  )
)
}
```

### 1. Read in the datasets
<br>
```{r,warning=FALSE,message=FALSE}

hits_df <- read_csv("ML Data Scientist Case Study Data.csv")[1:50000,]

```

### 2. Clean the variables a bit
<br>
```{r,warning=FALSE,message=FALSE}

hits_df <- hits_df %>%
  separate(`row_num;locale;day_of_week;hour_of_day;agent_id;entry_page;path_id_set;traffic_type;session_durantion;hits`,into = c("A","B","C"),sep = '\\"')

hits_df <- hits_df %>%
  separate(A,into = c("row_num","locale","day_of_week","hour_of_day","agent_id","entry_page","path_id_set","traffic_type","session_duration","hits"),remove = T)

hits_df <- hits_df %>% 
  mutate(path_id_set = ifelse(path_id_set == "",B,path_id_set)) %>% 
  select(-B)


hits_df <- hits_df %>%
  separate(C,into = c("D","A","B","C2"),sep = ';')

hits_df$traffic_type <- ifelse(is.na(hits_df$traffic_type),hits_df$A,hits_df$traffic_type)
hits_df$session_duration <- ifelse(is.na(hits_df$session_duration),hits_df$B,hits_df$session_duration)

hits_df$hits <- ifelse(is.na(hits_df$hits),hits_df$C2,hits_df$hits)

hits_df <- hits_df %>% select(-D:-C2)

hits_df <- hits_df %>% mutate(mashida = ifelse(is.na(hits),1,0))

hits_df <- hits_df %>% 
  mutate(hits = ifelse(mashida==1,session_duration, hits),
        session_duration = ifelse(mashida==1,traffic_type, session_duration),
        traffic_type = ifelse(mashida==1,path_id_set, traffic_type),
        path_id_set = ifelse(mashida==1,0, path_id_set))


hits_df$hits <- ifelse(hits_df$hits %in% grep("N",hits_df$hits,value = T, ignore.case = T),NA,hits_df$hits)
hits_df$session_duration <- ifelse(hits_df$session_duration %in% grep("N",hits_df$session_duration,value = T, ignore.case = T),NA,hits_df$session_duration)

hits_df$hits <- as.numeric(hits_df$hits)
hits_df$session_duration <- as.numeric(hits_df$session_duration)

hits_df$mashida <- NULL

hits_df <- data.frame(hits_df)



```


###3. Carry out some feature engineering.

We are going to create other ariables, based on the ones we have, which may be better predictors of the number of hits.

#### 3.1 Generate two variables, one showing whether the day is a weekday or not, another showing whether the day is at the beginning or at the end of the week.

##### 3.1.1 Day of Week variable
<br>
```{r,warning=FALSE,message=FALSE,results='asis'}
## First factor the current weekday variable

hits_df <- hits_df %>% 
  mutate(day_of_week = fct_relevel(day_of_week,"Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday"))

## Plot the mean number of hits for each of the categories in the three variables.

summ_tab <- hits_df %>% 
  dplyr::group_by(day_of_week) %>% 
  summarise(avg_hits = round(mean(hits, na.rm = TRUE),2)) 

## Print the table

pr_func(summ_tab,cnames = c("Day of the week","Average No. of Hits") )


## Graph
summ_graph <- ggplot(summ_tab ,aes(x=day_of_week,y=avg_hits,group = 1))+
              geom_line(stat = "identity",color = trivago_yellow)+
              geom_text(aes(label = avg_hits),vjust = -0.25, size = 5)+
              trivago_theme+
              labs(title = "Average Number of Hits Per Session",x="Day of Week",
                   y="Average Number of Hits")
summ_graph


```

##### 3.1.2 Weekend variable
<br>

```{r,warning=FALSE,message=FALSE,results='asis'}

## Generate the weekend variable
hits_df <- hits_df %>% 
  mutate(weekend = ifelse(day_of_week=="Sunday"|day_of_week=="Saturday","Weekend","Weekday"),
         weekend = fct_relevel(weekend,"Weekday","Weekend"))

## Plot the mean number of hits for each of the categories in the three variables.

summ_tab <- hits_df %>% 
  dplyr::group_by(weekend) %>% 
  summarise(avg_hits = round(mean(hits, na.rm = TRUE),2)) 

## Print the table

pr_func(summ_tab,cnames = c("Weekend","Average No. of Hits"))


## Graph
summ_graph <- ggplot(summ_tab ,aes(x=weekend,y=avg_hits))+
              geom_bar(stat = "identity",fill = trivago_yellow)+
              geom_text(aes(label = avg_hits),vjust = -0.25, size = 5)+
              trivago_theme+
              labs(title = "Average Number of Hits Per Session",x="",
                   y="Average Number of Hits")+
  ylim(c(0,20))
summ_graph

```

##### 3.1.3 Week Position variable
<br>

```{r,warning=FALSE,message=FALSE,results='asis'}

## Generate the week_pos variable that indicates whether the day is at the beginning or at the end of the week.
hits_df <- hits_df %>% 
  mutate(week_pos = ifelse(day_of_week=="Sunday"|day_of_week=="Monday"|day_of_week=="Tuesday"|day_of_week=="Wednesday","Beginning","End"),
         week_pos = fct_relevel(week_pos,"Beginning","End"))

## Plot the mean number of hits for each of the categories in the three variables.

summ_tab <- hits_df %>% 
  dplyr::group_by(week_pos) %>% 
  summarise(avg_hits = round(mean(hits, na.rm = TRUE),2)) 

## Print the table

pr_func(summ_tab,cnames = c("week_pos","Average No. of Hits") )



## Graph
summ_graph <- ggplot(summ_tab ,aes(x=week_pos,y=avg_hits))+
              geom_bar(stat = "identity",fill = trivago_yellow)+
              geom_text(aes(label = avg_hits),vjust = -0.25, size = 5)+
              trivago_theme+
              labs(title = "Average Number of Hits Per Session",x="",
                   y="Average Number of Hits")+
  ylim(c(0,20))
summ_graph

```


##### 3.1.4 Check if there is a relationship between these variables and the number of hits
<br>
```{r,warning=FALSE,message=FALSE,comment=NA}

summary(aov(hits~day_of_week, data = hits_df))
summary(aov(hits~weekend, data = hits_df))
summary(aov(hits~week_pos, data = hits_df))
```

#### 3.2 Hour of day

##### 3.2.1 Continous variable

```{r,warning=FALSE,message=FALSE,results='asis'}

## Plot the mean number of hits per hour.

summ_tab <- hits_df %>% 
  mutate(hour_of_day = as.numeric(hour_of_day)) %>% 
  group_by(hour_of_day ) %>% 
  summarise(avg_hits = round(mean(hits, na.rm = TRUE),2)) 


## Graph
summ_graph <- ggplot(summ_tab ,aes(x=as.factor(hour_of_day),y=avg_hits,group = 1))+
              geom_line(stat = "identity",color = trivago_yellow)+
              geom_text(aes(label = avg_hits),vjust = -0.25, size = 4)+
              trivago_theme+
              labs(title = "Average Number of Hits Per Session",x="Hour",
                   y="Average Number of Hits")
summ_graph



```

##### 3.2.2 Re-bucket the variable into categories

<br>
```{r,warning=FALSE,message=FALSE,results='asis'}

hits_df <- hits_df %>% 
  mutate(hour_of_day = as.numeric(hour_of_day)) %>% 
  mutate(hour_of_day2 = ifelse(hour_of_day >=0 & hour_of_day <=5,"Early Morning",
                        ifelse(hour_of_day >=6 & hour_of_day <=11,"Morning",
                        ifelse(hour_of_day >=12 & hour_of_day <=18,"Afternoon",
                        ifelse(hour_of_day >=19 & hour_of_day <=23,"Evening","")))))

hits_df <- hits_df %>% 
  mutate(hour_of_day2 = fct_relevel(hour_of_day2, "Early Morning","Morning","Afternoon","Evening")) 

summ_tab <- hits_df %>% 
  dplyr::group_by(hour_of_day2) %>% 
  summarise(avg_hits = round(mean(hits, na.rm = TRUE),2)) 

## Print the table

pr_func(summ_tab,cnames = c("Time of Day","Average No. of Hits") )



## Graph

summ_graph <- ggplot(summ_tab ,aes(x=as.factor(hour_of_day2),y=avg_hits,group = 1))+
              geom_line(stat = "identity",color = trivago_yellow)+
              geom_text(aes(label = avg_hits),vjust = -0.25, size = 4)+
              trivago_theme+
              labs(title = "Average Number of Hits Per Session",x="Time of Day",
                   y="Average Number of Hits")
summ_graph


```

##### 3.2.3 Testing if there is a correlation between hour of the day and the number of hits
<br>

```{r,warning=FALSE,message=FALSE,comment=NA}

cor.test(hits_df$hour_of_day, hits_df$hits) ## As the day progresses, the number of hits significantly decreases.



```

<br>
```{r,warning=FALSE,message=FALSE,results='asis'}
## Graph

summ_graph <- ggplot(hits_df ,aes(x=as.factor(hour_of_day),y=hits))+
              geom_point(color = trivago_yellow)+
              geom_smooth()+
              trivago_theme+
              #theme(axis.text.x = element_text(angle=45))+
              labs(title = "Number of Hits \n by\n Hour",x="Hour of Day",
                   y="Number of Hits")
summ_graph

```

#### 3.3 Path_id_set

##### 3.3.1 Number of pages visited
<br>

```{r,warning=FALSE,message=FALSE,comment=NA}
loc <- hits_df%>% 
  select(row_num,path_id_set) %>% 
  separate_rows(path_id_set,sep = ";") %>% 
  mutate(path_id_set = as.numeric(path_id_set)) %>% 
  group_by(row_num) %>% 
  mutate(counter = seq_along(path_id_set)) %>% 
  mutate(path_first0 = ifelse(path_id_set == 0 & counter == 1, "Yes","No"),
         path_first0 = ifelse(any(path_first0=="Yes"),"Yes","No"))
  


num_locs <- loc%>% 
  group_by(row_num) %>% 
  count() %>% 
  rename(num_locs = n)

first0 <- loc %>% 
  distinct(row_num,path_first0)

hits_df2 <-Reduce(function(x,y) merge(x,y) ,list(hits_df,num_locs,first0))

hits_df2 <- hits_df2 %>% 
  mutate(path_first0 = fct_relevel(path_first0,"No","Yes"))

rm(hits_df)
rm(loc)
rm(first0)
rm(num_locs)

summ_tab <- hits_df2 %>% 
  dplyr::group_by(num_locs) %>% 
  summarise(avg_hits = round(mean(hits, na.rm = TRUE),2)) 

## Print the table

pr_func(summ_tab,cnames = c("Number of pages visited","Average No. of Hits") )



## Graph

summ_graph <- ggplot(summ_tab ,aes(x=num_locs,y=avg_hits,group = 1))+
              geom_line(stat = "identity",color = trivago_yellow)+
              geom_text(aes(label = avg_hits),vjust = -0.25, size = 4)+
              trivago_theme+
              labs(title = "Average Number of Hits Per Session",x="Number of pages visited",
                   y="Average Number of Hits")
summ_graph

```


##### 3.3.2 First page visited being 0

```{r,warning=FALSE,message=FALSE,results='asis'}

summ_tab <- hits_df2 %>% 
  dplyr::group_by(path_first0) %>% 
  summarise(avg_hits = round(mean(hits, na.rm = TRUE),2)) 

## Print the table

pr_func(summ_tab,cnames = c("path_first0","Average No. of Hits") )



## Graph

summ_graph <- ggplot(summ_tab ,aes(x=as.factor(path_first0),y=avg_hits))+
              geom_bar(stat = "identity",fill = trivago_yellow)+
              geom_text(aes(label = avg_hits),vjust = -0.25, size = 5)+
              trivago_theme+
              labs(title = "Average Number of Hits Per Session",x="path_first0",
                   y="Average Number of Hits")
summ_graph

```




##### 3.3.3 Testing relationships between the two new variables and the hits variable

<br>
```{r,warning=FALSE,message=FALSE,comment=NA}

cor.test(hits_df2$num_locs, hits_df2$hits) ## Strong positive correlation. The higher the number of pages visited, the higher the higher the number of hits

summary(aov(hits~path_first0, data = hits_df2)) ## The number of hits is significantly higher for those whose first page is 0. 

```

#### 3.4 Locale

<br>
```{r,warning=FALSE,message=FALSE,comment=NA}

hits_df2 <- hits_df2 %>% 
  mutate(locale = fct_relevel(locale,"L1","L2","L3","L4","L5","L6"))

summary(aov(hits~locale, data = hits_df2))

```


```{r,warning=FALSE,message=FALSE,results='asis'}

summ_tab <- hits_df2 %>% 
  dplyr::group_by(locale) %>% 
  summarise(avg_hits = round(mean(hits, na.rm = TRUE),2)) 

## Print the table

pr_func(summ_tab,cnames = c("locale","Average No. of Hits") )



## Graph

summ_graph <- ggplot(summ_tab ,aes(x=as.factor(locale),y=avg_hits))+
              geom_bar(stat = "identity",fill = trivago_yellow)+
              geom_text(aes(label = avg_hits),vjust = -0.25, size = 5)+
              trivago_theme+
              labs(title = "Average Number of Hits Per Session",x="locale",
                   y="Average Number of Hits")
summ_graph
```


#### 3.5 Agent ID

<br>
```{r,warning=FALSE,message=FALSE,comment=NA}
hits_df2 <- hits_df2 %>% 
  mutate(agent_id = as.factor(as.numeric(agent_id)))

summary(aov(hits~agent_id, data = hits_df2))
```


```{r,warning=FALSE,message=FALSE,results='asis'}

summ_tab <- hits_df2 %>% 
  dplyr::group_by(agent_id) %>% 
  summarise(avg_hits = round(mean(hits, na.rm = TRUE),2)) 

## Print the table

pr_func(summ_tab,cnames = c("agent_id","Average No. of Hits") )



## Graph

summ_graph <- ggplot(summ_tab ,aes(x=as.factor(agent_id),y=avg_hits))+
              geom_bar(stat = "identity",fill = trivago_yellow)+
              geom_text(aes(label = avg_hits),vjust = -0.25, size = 5)+
              trivago_theme+
              labs(title = "Average Number of Hits Per Session",x="Agent ID",
                   y="Average Number of Hits")
summ_graph
```

#### 3.6 Entry Page

<br>
```{r,warning=FALSE,message=FALSE,comment=NA}

hits_df2 <- hits_df2 %>% 
  mutate(entry_page = as.factor(as.numeric(entry_page)))

summary(aov(hits~entry_page, data = hits_df2))

```


```{r,warning=FALSE,message=FALSE,results='asis'}

summ_tab <- hits_df2 %>% 
  dplyr::group_by(entry_page) %>% 
  summarise(avg_hits = round(mean(hits, na.rm = TRUE),2)) %>% 
  arrange(desc(avg_hits))

## Print the table

pr_func(summ_tab,cnames = c("Entry Page","Average No. of Hits"))

```


#### 3.7 Traffic Type

<br>
```{r,warning=FALSE,message=FALSE,comment=NA}

hits_df2 <- hits_df2 %>% 
  mutate(traffic_type = fct_relevel(traffic_type,"1","2","3","4","6","7","10"))

summary(aov(hits~traffic_type, data = hits_df2))

```


```{r,warning=FALSE,message=FALSE,results='asis'}

summ_tab <- hits_df2 %>% 
  dplyr::group_by(traffic_type) %>% 
  summarise(avg_hits = round(mean(hits, na.rm = TRUE),2)) 

## Print the table

pr_func(summ_tab,cnames = c("Traffic Type","Average No. of Hits") )



## Graph

summ_graph <- ggplot(summ_tab ,aes(x=as.factor(traffic_type),y=avg_hits))+
              geom_bar(stat = "identity",fill = trivago_yellow)+
              geom_text(aes(label = avg_hits),vjust = -0.25, size = 5)+
              trivago_theme+
              labs(title = "Average Number of Hits Per Session",x="Traffic Type",
                   y="Average Number of Hits")
summ_graph

```

#### 3.8 Session Duration
<br>
```{r,warning=FALSE,message=FALSE,comment=NA}

hits_df2 <- hits_df2 %>% 
  mutate(session_duration = as.numeric(session_duration)) 

cor.test(hits_df2$session_duration, hits_df2$hits)# There is a strong positive correlation between  session duration and number of hits 
```

<br>
```{r,warning=FALSE,message=FALSE,results='asis'}
## Graph

summ_graph <- ggplot(hits_df2 ,aes(x=session_duration,y=hits))+
              geom_point(color = trivago_yellow)+
              geom_smooth()+
              trivago_theme+
              labs(title = "Number of Hits \n by\n Session Duration",x="Session Duration",
                   y="Number of Hits")+
              xlim(c(0,200))+ylim(c(0,50))
summ_graph

```

### 4. Run an outliers test to drop outliers in the session duration variable 
<br>
```{r,warning=FALSE,message=FALSE,results='asis'}
## Create a box plot to identify outliers.

boxplot_graph <- ggplot(hits_df2 ,aes(y=hits))+
              geom_boxplot(color = trivago_yellow)+
              trivago_theme+
              labs(title = "",
                   y="")
boxplot_graph

outliers <- boxplot(hits_df2$hits, plot=FALSE)$out

hits_df2 <- hits_df2 %>% 
  filter(!session_duration %in% outliers)

```


### 5. Select the variables to be used in fitting the predictive model

<br>
```{r,warning=FALSE,message=FALSE,results='asis'}

reg_df <- hits_df2 %>% 
  select(locale,agent_id, entry_page,weekend,week_pos,hour_of_day2,num_locs,path_first0,  traffic_type, session_duration,hits)

```

### 6. Split the data into train and test. The test data will be the one with missing values in the hits variable.

<br>
```{r,warning=FALSE,message=FALSE,results='asis'}

train <- reg_df %>% filter(!is.na(hits))
train<-train[complete.cases(train),]

test <- reg_df %>% filter(is.na(hits))

```

### 7. Fit a Random Forest model to identify the variables that are highly important in predicting the number of hits.
<br>

```{r,warning=FALSE,message=FALSE,results='asis'}

# Fitting model
forest <- randomForest(hits ~ ., data=train %>% select(-entry_page),ntree=500)

## Importance of variables
importance_dat = data.frame(forest$importance)
importance_dat$variables <- row.names(importance_dat)
importance_dat <-importance_dat %>%select(variables, MeanDecreaseGini)
importance_dat = arrange(importance_dat,desc(MeanDecreaseGini))


```

