---
title: <center><strong>Trivago Case Study</strong></center>
subtitle: <center><strong><h2>Challenge 428 - Building a predictive 'hits' model:Part 2</h2></strong></center>
author: <center><strong><h1>[Shelmith Nyagathiri Kariuki](https://github.com/Shelmith-Kariuki)</h1></strong></center>
date: <center><strong><h2>January 20, 2020</h2></strong></center>
output:
  html_document:
    toc: yes
    toc_depth: '6'
    code_folding: hide
---
Trivago is an online company that meets the growing demand for independent travel information. It offers an extensive hotel meta search to travellers. 

Trivago hosts different case studies on their website, that job applicants can work on.

Details about this particular can be found can be found [here](https://jobs.zealpath.com/m/case/detail/428). 

The variables of the dataset are defined below:

+ row_num: a number uniquely identifying each row.

+ locale: the platform of the session.

+ day_of_week: Mon-Fri, the day of the week of the session.

+ hour_of_day: 00-23, the hour of the day of the session.

+ agent_id: the device used for the session.

+ entry_page: describes the landing page of the session.

+ path_id_set: shows all the locations that were visited during the session.

+ traffic_type: indicates the channel the user cane through eg. search engine, email, ...

+ session_duration: the duration in seconds of the session.

+ hits: the number of interactions with the trivago page during the session.


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE,fig.width = 12)

```

### 0. Install the libraries required
<br>
```{r,warning=FALSE, message=FALSE}

## create a vector of packages to be installed
pkgs <- c("tidyverse","data.table","DT","lubridate","ggthemes",
          "randomForest","e1071","rpart","pROC","caret")

## Check if there are packages you want to load, that are not already installed. 
miss_pkgs <- pkgs[!pkgs %in% installed.packages()[,1]] 

## Installing the missing packages
if(length(miss_pkgs)>0){
  install.packages(miss_pkgs)
}
## Loading all the packages
invisible(lapply(pkgs,library,character.only=TRUE))

## Remove the objects that are no longer required
rm(miss_pkgs)
rm(pkgs)

```


<br>
```{r,warning=FALSE, message=FALSE}
### Setting the plot theme
trivago_theme<- theme_hc()+ theme(legend.position = "right", 
                 legend.direction = "vertical",
                 legend.title = element_blank(),
                 plot.title = element_text( size = rel(1.6), hjust = 0.5),
                plot.subtitle = element_text(size = rel(1.5), hjust = 0.5),
                 #axis.text = element_text( size = rel(1.5)),
                 axis.text.x = element_text(size =rel(1.5),angle = 0),
                 axis.text.y = element_text(size =rel(1.5),angle = 0),
                 axis.title = element_text( size = rel(1.55)),
                axis.line.x = element_line(size = 1.5, colour = "#c94a38"),
                 panel.background = element_rect(fill = NA))

### Colours that will be used for the plots 
trivago_blue = "#007faf"
trivago_orange = "#c94a38"
trivago_yellow = "#f48f00"

## Avoidance of scientific numbers
options(scipen = 999)

## Printing function
pr_func<-function(data,cnames){
datatable(data,colnames = cnames,
  extensions = 'Buttons', options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'print')
  )
)
}
```

### 1. Read in the datasets
<br>
```{r,warning=FALSE,message=FALSE}

hits_df <- readRDS("ML Data Scientist Case Study Data_Clean.rds")

```


### 1. Run an outliers test to drop outliers in the session duration variable 
<br>
```{r,warning=FALSE,message=FALSE,results='asis',fig.width=12}
## Create a box plot to identify outliers.

boxplot_graph <- ggplot(hits_df ,aes(y=hits))+
              geom_boxplot(color = trivago_yellow,outlier.colour = trivago_blue)+
              trivago_theme+
              labs(title = "Distribution of Hits (with outliers)",
                   y="Hits")
boxplot_graph

outliers <- boxplot(hits_df$hits, plot=FALSE)$out


boxplot_graph <- ggplot(hits_df %>% filter(hits < min(outliers)) ,aes(y=hits))+
              geom_boxplot(color = trivago_yellow,outlier.colour = trivago_blue)+
              trivago_theme+
              labs(title = "Distribution of Hits (without outliers)",
                   y="Hits")
boxplot_graph



hits_df <- hits_df %>% 
  filter(!session_duration %in% outliers)

rm(boxplot_graph)
```


### 2. Select the variables to be used in fitting the predictive model

<br>
```{r,warning=FALSE,message=FALSE,results='asis',fig.width=12}

reg_df <- hits_df %>% 
  select(row_num,locale,agent_id, entry_page,weekend,week_pos,hour_of_day2,num_locs,path_first0,  traffic_type, session_duration,hits)

```

### 3. Split the data into train and test. The test data will be the one with missing values in the hits variable.

<br>
```{r,warning=FALSE,message=FALSE,results='asis',fig.width=12}

train <- reg_df %>% filter(!is.na(hits))
train<-train[complete.cases(train),]
# train <- train[1:(nrow(train)/10),]

test <- reg_df %>% filter(is.na(hits))
# test <- test[1:(nrow(test)/10),]
```

### 4. Fit a Random Forest model to identify the variables that are highly important in predicting the number of hits.
<br>

```{r,warning=FALSE,message=FALSE,results='asis',fig.width=12}
# rm(hits_df2)
# rm(reg_df)
# Fitting model
forest <- randomForest(hits ~ ., data=train%>% select(-row_num,-entry_page),ntree=500)
varImpPlot(forest,main = "Variable Importance in predicting number of hits")

## Importance of variables
## More useful variables achieve higher increases in node purities
importance_dat = data.frame(forest$importance)
importance_dat$variables <- row.names(importance_dat)
importance_dat <-importance_dat %>%select(variables, IncNodePurity)
importance_dat = arrange(importance_dat,desc(IncNodePurity))

#Predict Output 
train_pred= predict(forest,newdata = train,type = 'response')

## Root Mean Squared error
rmse <- RMSE(pred = train_pred,obs = train$hits)



test_pred <- predict(forest,newdata=test %>% select(-hits),type = 'response')

test2 <- test 
test2$hits <- floor(test_pred)
test2 <- test2 %>% select(row_num,hits)
write.csv(test2,"RandomForest_results_full.csv")
```

### 5. Fit the prediction model using an SVM model.
<br>
```{r,warning=FALSE,message=FALSE,results='asis',fig.width=12}
# Fit the model
svm_model <- svm(hits ~ .,
                 data=train%>% select(-row_num,-entry_page),
                 kernel = "radial",
                 type = "nu-regression")
svm_model

## Predict Output 
train_pred= predict(svm_model,newdata=train ,type = 'response')
y<-train[,"hits"]

## Root Mean Squared error
rmse <- RMSE(pred = train_pred,obs = y) #30.53476 with 5000 

test_pred <- predict(svm_model,newdata=test %>% select(-hits),type = 'response')

test2 <- test %>% filter(!is.na(session_duration))
test2$hits <- floor(test_pred)
test2 <- test2 %>% select(row_num,hits)
write.csv(test2,"SVM_results_full.csv")

```


### 6. Fit the prediction model using a neural network model.

<br>

```{r,warning=FALSE,message=FALSE,results='asis',fig.width=12}
# Fit the model
library(nnet)

nnets_model <- nnet(hits ~ .,MaxNWts=5000,
                 data=train[1:10000,]%>% select(-entry_page), size=20)

#import the function from Github
library(devtools)
source_url('https://gist.githubusercontent.com/fawda123/7471137/raw/466c1474d0a505ff044412703516c34f1a4684a5/nnet_plot_update.r')
 
#plot the model
plot.nnet(nnets_model)


## Predict Output 
train_pred= predict(nnets_model,newdata = train,type = 'raw')

## Root Mean Squared error
rmse <- RMSE(pred = train_pred,obs = y) 

test_pred <- predict(svm_model,newdata=test %>% select(-hits),type = 'response')

test2 <- test %>% filter(!is.na(session_duration))
test2$hits <- floor(test_pred)
test2 <- test2 %>% select(row_num,hits)
write.csv(test2,"SVM_results.csv")
```
