---
author: "Shelmith Kariuki"
title: "PRS (Self Development)"
Task Number: "Task Number 1 (Reshaping Data)"
date: "7/23/2018"
output: html_document
---

#####i). Setting up the working directory
```{r setup, include=FALSE}
#-----------------------------------------------------
# 0. Set working environment
#-----------------------------------------------------
# drop all objects from memory
rm(list = ls())

# disable scientific notation
options(scipen = 999)

# set working directory
basedir <- normalizePath("/Users/busara/Documents/Github/R/")
# knitr::opts_knit$set(root.dir = basedir, base.dir = basedir)
# knitr::opts_chunk$set(echo = TRUE)
```

#####ii). Loading the required packages
```{r}
# load/install required packages
pkgs<-c("dplyr", "ggplot2","tidyr","tidyverse")


miss_pkgs <- pkgs[!pkgs %in% installed.packages()[,1]] # vector of missing packages

# install the missing packages
if(length(miss_pkgs)>0){
  install.packages(miss_pkgs)
}

# Loading all the packages
invisible(lapply(pkgs,library,character.only=TRUE))
```

#####1. Read in the data

```{r}
SalesData<-read.csv("SalesData.csv")
SalesData<-SalesData[,-1]
```

#####2. Reshape the data, from wide to long, to contain the following columns: County, Year, Quarters and Sales.

```{r}
SalesData_long<-SalesData%>%
  gather("Quarters","Sales",-Year)
```

#####3. Calculate the average yearly sales.

```{r}
sales_mean<-SalesData_long%>%
  group_by(Year)%>%
  summarise(Sales = mean(Sales))
```

#####4. Draw a neatly labelled line plot, that shows the average sales for each year.

```{r}
mean_sales1<-ggplot(data = sales_mean,
                    aes(x = as.factor(Year), y=Sales,group = 1))+
  geom_line(color = "red",size =1.5)+
  theme(legend.position = "none",
        legend.title = element_blank(),
        axis.line=element_blank(),
        plot.title = element_text(size = rel(1.5), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.4), hjust = 0.5),
        axis.text = element_text(size = 10),
        axis.text.x=element_text(vjust = 1,angle = 30,hjust = 0.9),
        axis.title = element_text(size =12),
        panel.background = element_rect(fill = NA))+
  labs(title ="Average Sales Values per Year", x = "Year", y = "Average Sales")
mean_sales1
```

#####5. Draw a neatly labelled line plot, that shows the quarterly sales for each year.

```{r}
mean_sales2<-ggplot(data = SalesData_long,
                    aes(x = as.factor(Year), y=Sales,group = Quarters,color =Quarters))+
  geom_point()+
  geom_line(size =1.5)+
  theme(legend.position = "right",
        legend.title = element_blank(),
        axis.line=element_blank(),
        plot.title = element_text(size = rel(1.5), hjust = 0.5),
        plot.subtitle = element_text(size = rel(1.4), hjust = 0.5),
        axis.text = element_text(size = 10),
        axis.text.x=element_text(vjust = 1,angle = 30,hjust = 0.9),
        axis.title = element_text(size =12),
        panel.background = element_rect(fill = NA))+
  scale_fill_brewer(palette = "Spectral")+
  labs(title ="Quarterly Sales Values per Year", x = "Year", y = "Sales")
mean_sales2

```

