---
title: <strong><center>Beyond ordinary mutate(),summarise() functions</strong></center>
author: <strong><center>Shelmith</strong></center>
date: <strong><center>1/14/2020</strong></center>
output: html_document
---

I have comfortably used the ordinary mutate(),summarise() functions for almost 3 years now, but I thought it is high time I tried the different variations i.e mutate_all(), mutate_if(), mutate_at(),summarise_all(), summarise_if(), summarise_at(). I will also share some other tricks that I came acroos, while learning these functions.

Motivation of preparing this R Markdown document is derived from Suzan Bert's article found [here](https://suzan.rbind.io/2018/02/dplyr-tutorial-2/).

We will use the "Wafanyikazi" dataset that can be found on the [rChambua](https://github.com/Shelmith-Kariuki/rChambua) pakage. This dataset is totally hypothetical. I generated it after wanting a dataset that I could play with all the time.

### 0. Load the packages that we will require
<br> 

```{r,warning=FALSE, message=FALSE}

## Create a vector of packages that we will need
pkgs <- c("dplyr")

## Check if there are variables you want to load, that are not already installed. 
miss_pkgs <- pkgs[!pkgs %in% installed.packages()[,1]] 

## Installing the missing packages
if(length(miss_pkgs)>0){
  install.packages(miss_pkgs)
}

## Loading all the packages
invisible(lapply(pkgs,library,character.only=TRUE))

## Remove the objects that are no longer required
rm(miss_pkgs)
rm(pkgs)

## Install rChambua from Github, and load it
#install.packages("rChambua")
library("rChambua")

```

### 1. Read in the dataset, and inspect it using glimpse()
<br> 

```{r,warning=FALSE, message=FALSE}
df <- wafanyikazi

glimpse(df)
```

### 2. mutate()

This is usually the norm for me. 

Let us generate a Swahili version of Gender, call it Jinsia, which should have the values "Mme" inplace of "Male", and "Mke" inplace of "Female".

<br> 
```{r,warning=FALSE, message=FALSE}

df <- df %>% 
  mutate(Jinsia = ifelse(Gender == "Female", "Mke", "Mme"))
glimpse(df)

head(df, 10)
```

### 3. mutate_all()

This function is used to implement a function on all the variables.

The syntax is written as:
mutate_all(inbuilt function) or
mutate_all(~(custom function with a "." representing each var)) or
mutate_all(funs(custom function with a "." representing each var)) 

Let us assume that we want to include an "_x" at the end of each value, in all the variables

<br>
```{r,warning=FALSE, message=FALSE}

df2 <- df %>% 
  mutate_all(~paste0(.,"_x"))

df2 <- df %>% 
  mutate_all(funs(paste0(.,"_x")))

head(df2, 10)

rm(df2)
```

Seems as though funs() is deprecated, when using mutate_all().

### 4. mutate_if()

Used to implement a function on some variables, if they meet a certain condition.

The syntax as written as: mutate_if(condition that the variables should meet, the function to be implemented).

Let us generate a dataset that contains all the numeric variables mutiplied by 100,
and all the factor/character variables converted to lower case.

<br> 
```{r,warning=FALSE, message=FALSE}

df2 <- df %>% 
  mutate_if(is.numeric, ~(.*100))

df2 <- df %>% 
  mutate_if(is.factor, tolower)

df2 <- df %>% 
  mutate_if(is.factor, ~(paste("New_",tolower(.))))

head(df2, 10)
rm(df2)           
```

### 5. mutate_at()

Used to implement a function on some selected variables

The syntax as written as: mutate_at(vars(variables), function to be implemented)

Generate a dataset that contains the variables age and income divided by 5.
<br>

```{r,warning=FALSE, message=FALSE}

df2 <- df %>% 
  mutate_at(vars(Age, Income), funs(div =./5))

head(df2, 10)
rm(df2) 

```

### 6. summarise_all()
<br>

```{r,warning=FALSE, message=FALSE}


  
```

### 7. summarise_if()

<br> 
```{r,warning=FALSE, message=FALSE}

tab_mean <- df %>% 
  group_by(Department) %>% 
  summarise_if(is.numeric, mean, na.rm = TRUE) %>% 
  mutate(metric = "Mean")
tab_mean

tab_median <- df %>% 
  group_by(Department) %>% 
  summarise_if(is.numeric, median, na.rm = TRUE) %>% 
  mutate(metric = "Median")
tab_median

tab <-bind_rows(tab_mean, tab_median)
tab

rm(tab_mean)
rm(tab_median)
rm(tab)


```

### 8. summarise_at()
<br> 

```{r,warning=FALSE, message=FALSE}

tab <- df %>% 
  summarise_at(vars(Age,Leave_Days), max)
tab
```

### 9. unite()

Used to combine two variables or more, into 1

<br> 
```{r,warning=FALSE, message=FALSE}
df <- df %>% 
  mutate_at(vars(Gender, County), as.character) %>% 
  unite(United, Gender, County, sep=":",remove = F)
head(df, 10)
```

### 10. Generating a dataset of counts, grouped by two variables, and arranging the counts by ascending / descending order,

<br> 
```{r,warning=FALSE, message=FALSE}
##  Before

tab <- df %>% 
  group_by(Gender, Department) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))
tab
##  Now

tab <- df %>% 
  count(Gender, Department, sort = T,name = "count")
tab
```


