---
title: <center><strong>"Zukademy Assignment"</strong></center>
date: <center>"April 19, 2019"</center>
output: html_document
---

```{r setup, include=FALSE}
name<-"insert your name here"
```

This markdown was produced by `r name` on `r Sys.Date()`

<center><h2>Introduction</h2></center>

These assignment covers all the tidyverse based topics covered so far. 

The data used in Q1-Q10 can be found here <https://zindi.africa/competitions/mobile-money-and-financial-inclusion-in-tanzania-challenge/data>

Download the **training.csv** dataset, and the **Variable_Codebook.docx** document.

### 1. Read in the dataset
<br>
```{r}
## write code here
```

### 2. Skim through the assignment, and load the libraries that you will require.

<br>
```{r}
## write code here
```

### 3. Recode all the variables that need to be recoded, based on Variable_Codebook document

<br>

```{r}
## write code here
### Hint: instead of gender reading 1, 2, the values should read Male, Female. Remember ifelse(condition, result if condition is True, result if condition is false)

### Q2: Gender

### Q3: Marital Status

### Q4: Highest level of education

### ...

### Mobile_money_classification



### These chunk will be long. Apologies in advance, but practice practice makes perfect

```

###4. Generate frequency tables and graphs, for all the categorical variables

####4.1 Gender

<br>
```{r}
## write code here
```

####4.2 Marital Status

<br>

```{r}
## write code here
```

####4.3 Highest level of education

<br>
```{r}
## write code here
```

#### And so on... Each chunk should have results (table and graph) for one variable.

<br>
```{r}
## write code here
```

###5 Which gender:

#### 5.1 Saves the most?
<br>
```{r}
## write code here
```

#### 5.2 Borrows the most?
<br>
```{r}
## write code here
```

#### 5.3 Uses insurance the most?
<br>
```{r}
## write code here
```


###6. Between those who save, and those who do not, which group has a higher literacy level? (both english and kiswahili)
<br>
```{r}
## write code here
```

###7. How many single females, aged 50 and below, have insurance and use mobile money only? How many of them received cash in the past 30 days
<br>
```{r}
## write code here
```

###8. What is the mean age of:

####8.1 Borrowers vs Non Borrowers
<br>
```{r}
## write code here
```

####8.2 Those who save, vs those who do not
<br>
```{r}
## write code here
```

####8.3 Those who can read only vs those who can write only, in english 

<br>
```{r}
## write code here
```

###9. How many married respondents, with no formal education, work for the government?
<br>
```{r}
## write code here
```


###10. Think of any other analysis that can be done using these data and add it here.

<br>
```{r}
## write code here
```



###11. The data generated below shows the quarterly perfomance of students, in zukademy

<br>
```{r}
## write code here
cols<-c("2011Q1","2011Q2","2011Q3","2011Q4",
     "2012Q1","2012Q2","2012Q3","2012Q4",
     "2013Q1","2013Q2","2013Q3","2013Q4",
     "2014Q1","2014Q2","2014Q3","2014Q4",
     "2015Q1","2015Q2","2015Q3","2015Q4")

rows<-c("Males","Females")

data<-data.frame(matrix(sample(100:500,length(cols)*length(rows),replace = F),ncol=length(cols), nrow = length(rows), byrow = T))

colnames(data) = cols
rownames(data)=rows

```


#### 11.1 Reshape the data, so that you get a dataset with three columns,Gender, Year, Quarter and Marks
<br>
```{r}
## write code here
```


#### 11.2 Calculate the average marks per gender, .per year
<br>
```{r}
## write code here
```

#### 11.3 Represent the results in 11.2 on a line graph
<br>
```{r}
## write code here
```

#### 11.4 Calculate the average marks per gender, per year and per quarter
<br>
```{r}
## write code here
```

#### 11.5 Calculate the total marks,for both genders, per quarter and per year
<br>
```{r}
## write code here
```


#### 11.6 Represent the results in 11.5 on a line graph.
<br>
```{r}
## write code here
```

#### 11.7 All those who score 300 and above, are taken for dinner at Kempinski. How many ladies got their Kempinski treat in 2013 and 2014?
<br>
```{r}
## write code here
```





